Figures to insert in text

### 2018-03-05 Chapters/sdm2.tex
Review of the SDM description

The data and address space belong to binary space and are represented by a sequence of bits, called bitstrings. The distance between two bitstrings is calculated using the Hamming distance. It is defined for two bitstrings of equal length as the number of positions at which the bits are different. For example, $00110_{b}$ and $01100_{b}$ are bitstrings of length 5 and their Hamming distance is 2. One has to be careful when thinking intuitively about distance in SDM because the Hamming distance does not have the same properties of, say, the Euclidean distance.

@@

The main change between Kanerva's read and Chada's read is that, in the former, a hard-location that has more bitstrings written has a greater weight in the decision of each bit. In the latter, all hard-locations have the same weight, because they can contribute to the sum with only one bitstring.

A member of my Master's committee, professor Paulo Murilo, has proposed a generalized reading operation (personal communication), which covers both Kanerva's and Chada's read --- and opens a new venue of potential discoveries. He proposed summing all hard-location counters raised to the power of $z$ while holding the original sign of the counter (positive or negative). Thus, Kanerva's read would be the same as $z=1$, while Chada's would be the same as $z=0$. Hence, we will here explore how SDM would behave with other values of $z$, such as 0.5, 2, and 3.

@@

In his book, Kanerva analyzed a specific situation with $n=1000$ ($N=2^{1000}$), 10 million hard-locations, an access-radius of 451 (within 1000 hard-locations in each circle) and 10 thousand writes of random bitstrings in the memory. As computer resources were very poor those days, Kanerva couldn't make a more generic analysis.

@@

Figure \ref{fig:crit-dist-10k-writes} compares the critical distance behavior under different scenarios.  This replicates our previous results \citet{Brogliato2011} and \citet{brogliato2014sparse} and is a first part of the process of framework validation, to which we throw our attention next.

>>>>

The data and address space belong to binary space and are represented by a sequence of bits, called bitstrings. The distance between two bitstrings is calculated using the Hamming distance. It is defined for two bitstrings of equal length as the number of positions at which the bits are different. For example, $00110_{b}$ and $01100_{b}$ are bitstrings of length 5 and their Hamming distance is 2. One has to be careful when thinking intuitively about distance in SDM because the Hamming distance does not have the same properties of, say, the Euclidean distance in 3 dimensions, i.e., both follow triangle inequality ($d(A, B) \le d(A, C) + d(B, C)$), which Euclidean distance in 3 dimensions may be interpreted as ``if A is close to B, and B is close to C, then A is also close to C'' --- $d(A, B) \le r \text{and} d(B, C) \le r \Rightarrow d(A, C) \le 2r$ ---, but in SDM, although the inequality is also valid, two bitstrings would be close when, for instance, $r = 430$, so $2r = 860$ would cover all other bitstrings. Hence, it makes no sense to say that $A$ is also close to $C$. This difference in intuition may trick even experienced researchers when analyzing some situations.

@@

The main difference between Kanerva's read and Chada's is that, in the former, a hard-location that has more bitstrings written has a greater weight in the decision of each bit. In the latter, all hard-locations have the same weight, because they can contribute to the sum with only one bitstring.

It is important to say that Chada's read came from \citet{anwar2003sparse} which gave a misguided description of the read operation. The original description is the following:

\begin{samepage}
\begin{quote}
Similar to writing, retrieving from SDM involves the same concept of access sphere --- you read all the hardlocations within the access sphere of location Y, pool the bit vectors read from all these hard locations and let each of the kth bits of those locations participate in a majority vote for the kth bit of Y. \hfill --- \citet{anwar2003sparse}
\end{quote}
\end{samepage}

This fact just highlihts how important it is to have a reference implementation that one may read the code to clarify one's understanding about the details of each operation.

\subsection{Generalized read operation}

A member of my Master's committee, professor Paulo Murilo, has proposed a generalized reading operation (personal communication), which covers both Kanerva's and Chada's read --- and opens a new venue of potential discoveries. He proposed summing all hard-location counters raised to the power of $z$ while holding the original sign of the counter (positive or negative). Thus, Kanerva's read would be the same as $z=1$, while Chada's would be the same as $z=0$. Hence, we will here explore how SDM would behave with other values of $z$, such as 0.5, 2, and 3. Mathematically, let $A$ be the set of the counters of the activated hardlocation, and $c_i$ be the counter of the $i$-th bit. Then,

$$
s_i = \sum_{c \in A} \frac{c_i}{|c_i|} |c_i|^z
$$

Finally, the $i$-th bit of the resulting bitstring is 1 if $s_i > 0$, or 0 if $s_i < 0$, or random if $s_i = 0$. Notice that when $z=1$, then $s_i = \sum_{c \in A} c_i$, which is the Kanerva's read; and when $z=0$, then $s_i = \frac{c_i}{|c_i|} = \text{sign}(c_i)$, which is the Chada's read.

@@

In his book, Kanerva analyzed a specific situation with $n=1000$ ($N=2^{1000}$), 1 million hard-locations $N'=1,000,000$, an access-radius of 451 (within 1,000 hard-locations in each circle) and 10 thousand writes of random bitstrings in the memory. As computer resources were very poor those days, Kanerva couldn't make a more generic analysis.

@@

Figure \ref{fig:crit-dist-10k-writes} compares the critical distance behavior under different scenarios.  This replicates our previous results \citep{Brogliato2011, brogliato2014sparse} and is a first part of the process of framework validation, to which we throw our attention next.
